EDA

### 2024-11-28 other/EDA.qmd
added more graphs

@@

# Create the bar chart
ggplot(avg_contract_size, aes(x = reorder(buyer, `Average Contract Size`), y = `Average Contract Size`)) +
  geom_bar(stat = "identity", fill = "darkgreen") +
  coord_flip() +
  scale_y_continuous(labels = dollar_format(prefix = "$")) +
  labs(x = "Government Department", y = "Average Contract Size", title = "Average Contract Size by Government Department (Top 10)") +
  theme_minimal()

@@

>>>>

```
#TODO Color code by buyer

@@

# Create the bar chart
ggplot(avg_contract_size, aes(x = reorder(buyer, `Average Contract Size`), y = `Average Contract Size`)) +
  geom_bar(stat = "identity", fill = "darkgreen") +
  coord_flip() +
  scale_y_continuous(labels = dollar_format(prefix = "$")) +
  labs(x = "Government Department", y = "Average Contract Size", title = "Average Contract Size by Government Department (Top 10)") +
  theme_minimal()
```

```{r}
# Convert award_date to Date type and extract the year
df$award_date <- as.Date(df$award_date, format = "%Y-%m-%d")
df$year <- format(df$award_date, "%Y")

# Filter for military spending (where buyer is "National Defence")
military_df <- df %>%
  filter(buyer == "National Defence")

# Summarize total spending per year
military_spending_per_year <- military_df %>%
  group_by(year) %>%
  summarize(total_spending = sum(amount, na.rm = TRUE))

# Create the bar chart
ggplot(data = military_spending_per_year, aes(x = year, y = total_spending)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Military Spending Per Year", x = "Year", y = "Total Spending") +
  theme_minimal() +
  scale_y_continuous(labels = dollar_format(prefix = "$")) +
  theme(plot.title = element_text(hjust = 0.5))

@@

```{r}

df$year <- format(df$award_date, "%Y")

# Filter for military spending (where buyer is "National Defence")
public_services_df <- df %>%
  filter(buyer == "Public Services and Procurement Canada")

# Summarize total spending per year
public_spending_per_year <- public_services_df %>%
  group_by(year) %>%
  summarize(total_spending = sum(amount, na.rm = TRUE))

# Create the bar chart
ggplot(data = public_spending_per_year, aes(x = year, y = total_spending)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Public Services Spending Per Year", x = "Year", y = "Total Spending") +
  theme_minimal() +
  scale_y_continuous(labels = dollar_format(prefix = "$")) +
  theme(plot.title = element_text(hjust = 0.5))

```{r}
df$year <- format(df$award_date, "%Y")

# Filter data for the specified buyers
selected_buyers <- c(
  "National Defence",
  "Public Services and Procurement Canada",
  "Shared Services Canada",
  "Public Health Agency of Canada",
  "Fisheries and Oceans Canada"
)

stacked_df <- df %>%
  filter(buyer %in% selected_buyers) %>%
  group_by(year, buyer) %>%
  summarize(total_spending = sum(amount, na.rm = TRUE)) %>%
  ungroup()

# Ensure the order of the stacks matches the specified order
stacked_df$buyer <- factor(stacked_df$buyer, levels = selected_buyers)

# Create the stacked bar chart
ggplot(data = stacked_df, aes(x = year, y = total_spending, fill = buyer)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(
    values = c("National Defence" = "steelblue",
               "Public Services and Procurement Canada" = "orange",
               "Shared Services Canada" = "green",
               "Public Health Agency of Canada" = "purple",
               "Fisheries and Oceans Canada" = "red")
  ) +
  labs(
    title = "Spending by Selected Buyers Per Year",
    x = "Year",
    y = "Total Spending",
    fill = "Buyer"
  ) +
  theme_minimal() +
  scale_y_continuous(labels = dollar_format(prefix = "$")) +
  theme(plot.title = element_text(hjust = 0.5))

```
```{r}
df$year <- format(df$award_date, "%Y")

# Filter data for the specified buyers
selected_buyers <- c(
  "National Defence",
  "Public Services and Procurement Canada",
  "Shared Services Canada",
  "Public Health Agency of Canada",
  "Fisheries and Oceans Canada"
)

grouped_df <- df %>%
  filter(buyer %in% selected_buyers) %>%
  group_by(year, buyer) %>%
  summarize(total_spending = sum(amount, na.rm = TRUE)) %>%
  ungroup()

# Ensure the order of the groups matches the specified order
grouped_df$buyer <- factor(grouped_df$buyer, levels = selected_buyers)

# Create the grouped bar chart
ggplot(data = grouped_df, aes(x = year, y = total_spending, fill = buyer)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
  scale_fill_manual(
    values = c("National Defence" = "steelblue",
               "Public Services and Procurement Canada" = "orange",
               "Shared Services Canada" = "green",
               "Public Health Agency of Canada" = "purple",
               "Fisheries and Oceans Canada" = "red")
  ) +
  labs(
    title = "Spending by Selected Buyers Per Year",
    x = "Year",
    y = "Total Spending",
    fill = "Buyer"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

```{r}

df$year <- format(df$award_date, "%Y")

# Define the selected buyers
selected_buyers <- c(
  "National Defence",
  "Public Services and Procurement Canada",
  "Shared Services Canada",
  "Public Health Agency of Canada",
  "Fisheries and Oceans Canada"
)

# Add a new column for buyer categories
df <- df %>%
  mutate(buyer_category = ifelse(buyer %in% selected_buyers, buyer, "Other"))

# Summarize total spending per year for each buyer category
grouped_df <- df %>%
  group_by(year, buyer_category) %>%
  summarize(total_spending = sum(amount, na.rm = TRUE)) %>%
  ungroup()

# Ensure the order of the groups matches the specified order, with "Other" last
grouped_df$buyer_category <- factor(
  grouped_df$buyer_category,
  levels = c(selected_buyers, "Other")
)

# Create the grouped bar chart
ggplot(data = grouped_df, aes(x = year, y = total_spending, fill = buyer_category)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
  scale_fill_manual(
    values = c(
      "National Defence" = "steelblue",
      "Public Services and Procurement Canada" = "orange",
      "Shared Services Canada" = "green",
      "Public Health Agency of Canada" = "purple",
      "Fisheries and Oceans Canada" = "red",
      "Other" = "gray"
    )
  ) +
  labs(
    title = "Spending by Selected and Other Buyers Per Year",
    x = "Year",
    y = "Total Spending",
    fill = "Buyer"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
